Automatic Identification and Annotation of Parallel Tasks in Structured Programs

### 2017-11-09 docs/ACM/paper.tex
Changes in the introduction.

Annotation systems have risen to a place of prominence as a simple and
effective way to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11} and Tareador~\cite{Ayguade15}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
The compiler-related literature contains a generous load of success stories
describing the use of annotation systems to parallelize programs.
Combined with modern accelerators, they have led to substantial performance
gains~\cite{Bertolli14,Mendonca17,Poesia17,Reyes12,Wienke12}.
Yet, even if convenient, the use of annotations is not straightforward, and
lacks supporting tools.

Annotations such as OpenMP or OpenACC still require developers to
worry about typical hassles of the parallel world, such as race conditions
or deadlocks.
Moreover, because they are used in tandem with imperative programming languages,
such troubles are only made worse by pointer aliasing.
There exist tools that insert automatic annotations in
programs~\cite{Mendonca16,Pingali11,Nugteren14}.
All these technologies explore data-parallelism -- the possibility of running
the same computation independently on different data.
Task parallelism remains still uncharted land, in what concerns automatic
annotation.
Such fact is unfortunate, because much of the power of current annotation
systems lays on their ability to create tasks \ed{We need to find references!}.
This ability brings said systems closer to irregular programs such as those
that implement graphs and worklist algorithms \`{a} la
Pingali~\cite{Ben-Nun17,Kulkarni11,Pingali11}.
The goal of this work is to address this omission.

This paper describes \Taskminer, the first compiler that mines task parallelism
from programs.
To fulfill this goal \Taskminer{} solves three challenges.
First, it find program regions, e.g., loops or functions, that can be
effectively mapped onto tasks (Section~\ref{sub:region}).
Second, it determines symbolic bounds to all the memory blocks accessed within
those regions (Section~\ref{sub:symb}).
Third, it maps such information back into source code, producing annotations
that programmers can read (Section~\ref{sub:ir}).
The goal of this paper is to describe the design and the implementation of these
three technologies, and to demonstrate their benefit.

@@

@@

\subsection{Find Symbolic Bounds of Arrays}
\label{sub:symb}

\subsection{From IR to Source Code}
\label{sub:ir}

@@

>>>>

Annotation systems have risen to a place of prominence as a simple and
effective way to write parallel programs.
Examples of such systems include OpenMP~\cite{JaegerCP15},
OpenACC~\cite{OpenACC20}, OpenHMPP~\cite{Andion14}, OpenMPC~\cite{Lee10},
OpenSs~\cite{MeenderinckJ11}, Cilk++~\cite{Leiserson09} and
Tareador~\cite{Ayguade15}.
Annotations work as a meta-language: they let developers grant parallel
semantics to syntax originally written to execute sequentially.
The compiler-related literature contains a generous load of success stories
describing the use of annotation systems to parallelize programs.
Combined with modern accelerators, they have led to substantial performance
gains~\cite{Bertolli14,Mendonca17,Poesia17,Reyes12,Wienke12}.
Yet, even if convenient, the use of annotations is not straightforward, and
lacks supporting tools.

Annotations such as OpenMP or OpenACC still require developers to
worry about typical hassles of the parallel world, such as race conditions
or deadlocks.
Moreover, because they are used in tandem with imperative programming languages,
such troubles are only made worse by pointer aliasing.
There exist tools that insert automatic annotations in
programs~\cite{Mendonca16,Pingali11,Nugteren14}.
All these technologies explore data-parallelism -- the possibility of running
the same computation independently on different data.
Task parallelism remains still uncharted land, in what concerns automatic
annotation.
Such fact is unfortunate, because much of the power of current annotation
systems lays on their ability to create tasks~\cite{Ayguade09}.
This ability brings said systems closer to irregular programs such as those
that implement graphs and worklist algorithms \`{a} la
Pingali~\cite{Ben-Nun17,Kulkarni11,Pingali11}.
The goal of this work is to address this omission.

This paper describes \Taskminer, a compiler that mines task parallelism
from programs.
To fulfill this goal \Taskminer{} solves many challenges.
First, it finds program regions, e.g., loops or functions, that can be
effectively mapped onto tasks (Section~\ref{sub:region}).
Second, it determines symbolic bounds to all the memory blocks accessed within
those regions (Section~\ref{sub:symb}).
Third, it maps such information back into source code, producing annotations
that programmers can read (Section~\ref{sub:ir}).
Fourth, it extracts parameters from the code to estimate the profitability of
tasks, given the cost to create them.
These parameters feed conditional checks, which, at runtime, enable or disable
the creation of tasks (Section~\ref{sub:profit}).
Finally, \Taskminer{} determines which program variables need to be replicated at
each new task, or shared among them (Section~\ref{sub:variance}).
This paper describes the design and the implementation of these technologies, and
demonstrate their benefit.

@@

% The ability to handle pointers through speculative parallelization.

% when applied to general-purpose integer-intensive applications that have complex control flow and excessive pointer accesses, traditional parallelization tech- niques become quite ineffective, as they need to conservatively ensure program correctness by synchronizing all potential dependences in the program

% For example in 473.astar if we ignore dependences that only occur in less than 20% of all iterations, we can parallelize loops that correspond to 96% of total execution.

@@

\subsection{Finding Symbolic Bounds of Arrays}
\label{sub:symb}

\subsection{Mapping IR onto Source Code}
\label{sub:ir}

\subsection{Estimating the Profitability of Tasks}
\label{sub:ir}

\subsection{Separating Private from Shared Variables}
\label{sub:profit}

\subsection{Mapping IR onto Source Code}
\label{sub:variance}

@@

% Talk about annotations vs libraries, e.g., Wool, Intel TBB. See: "A Comparison of some recent Task-based Parallel Programming Models"
